\documentclass[letterpaper]{article}
\usepackage[submission]{aaai2027}
\usepackage[hyphens]{url}
\usepackage{graphicx}
\urlstyle{rm}
\def\UrlFont{\rm}
\usepackage{natbib}
\usepackage{caption}
\frenchspacing
\usepackage{booktabs}
\pdfinfo{
/TemplateVersion (2027.1)
}
\setcounter{secnumdepth}{0}

\title{PaperToSkill: Turning Research Papers into Portable Agent Skills}
\author{
    Anonymous Authors
}
\affiliations{
    Anonymous Institution\\
    anonymous@example.com
}

\begin{document}

\maketitle

\begin{abstract}
Many LLM and agent papers introduce workflows that could help ordinary users and
agent builders, but those workflows are usually described for human reading
rather than direct reuse. Summaries compress the idea, but often discard the
procedural details, validation checks, assumptions, and failure branches needed
to apply the method. We introduce PaperToSkill, a paper-to-skill conversion
workflow that turns source-anchored paper notes into compact, human-editable
agent skills. Each skill records the paper's contribution, use conditions,
workflow, validation checks, failure cases, transfer notes, and source anchors.
In a four-paper benchmark over AI Scientist-v2, Reflexion, AIDE, and
Toolformer, PaperToSkill generates skills that score 20/20 on deterministic
structural rubrics, preserve more deterministic operational coverage than
generic-summary and abstract-only baselines, remain under a 1200-word
compactness budget, and substantially reduce deterministic input-token proxy
size relative to full extracted papers. These results support PaperToSkill as a
reproducible conversion layer from curated paper notes to portable skills. All
four live-transfer saved-response sets are collected and scored under a
deterministic output-contract evaluator; these saved response files are not
human semantic fidelity or downstream execution-outcome evidence.
\end{abstract}

\section{Introduction}

Using LLMs is becoming a practical skill for researchers, builders, and
non-expert users. At the same time, many of the best ways to use LLM agents are
introduced in papers whose methods are difficult to operationalize. A paper may
describe a search policy, memory loop, validation stage, or interface contract,
but a user who wants to reuse the contribution must translate prose into an
agent workflow by hand.

PaperToSkill studies a simple hypothesis: papers can be converted into compact,
editable skills that preserve enough procedural knowledge for agents and users
to reuse the main contribution. We use ``skill'' to mean a natural-language
operational artifact with an entry-point \texttt{SKILL.md}, similar to formats
used by agent harnesses that load task-specific instructions. Unlike a summary,
a skill is not only explanatory. It should specify when to use the method,
which inputs are required, which steps to execute, how to validate progress,
which failures to watch for, and how to adapt the instructions across
harnesses.

This paper takes an artifact-first view. PaperToSkill is not presented as a
fully automatic PDF understanding system. The current implementation converts
curated, source-anchored paper notes into skills and source maps. This narrow
scope is intentional: it lets us evaluate whether the conversion layer
preserves operational knowledge before claiming end-to-end PDF automation.

Our current contributions are: (1) a paper-to-skill schema for operationalizing
research contributions; (2) a deterministic extraction scaffold that emits
\texttt{SKILL.md} plus source maps; (3) deterministic extracted-text-to-note
scaffolds evaluated on Toolformer and AIDE; (4) a four-paper benchmark using AI
Scientist-v2, Reflexion, AIDE, and Toolformer; and (5) deterministic/offline
evaluations for structure, context coverage, compactness, source grounding, and
transfer readiness.

\section{Related Work}

PaperToSkill is motivated by automated research and agent workflow systems.
AI Scientist-v2 uses agentic tree search and staged experiment management for
automated scientific discovery~\cite{lu2025aiscientistv2}. AIDE frames ML
engineering as search over code solutions~\cite{aide2025}. Reflexion stores
verbal feedback in episodic memory to improve later attempts without weight
updates~\cite{shinn2023reflexion}. Toolformer shows how language models can
generate and filter tool-use examples for API calls~\cite{schick2023toolformer}.
Skill-library and interface work such as Voyager and SWE-agent also demonstrate
that reusable skills and agent-facing interfaces can determine whether a method
transfers beyond its original implementation~\cite{wang2023voyager,yang2024sweagent}.
Paper2Agent is the closest concurrent system: it converts papers and associated
codebases into MCP servers with tools, resources, prompts, tests, and an
interactive paper-agent layer~\cite{miao2025paper2agent}. AgenticSciML shows a
related use of specialized agents, structured debate, method memory, and
evolutionary search for scientific machine-learning discovery~\cite{jiang2026agenticsciml}.

These systems show that papers can become active computational artifacts.
PaperToSkill targets a different conversion layer: compact, human-editable
skills that preserve operational knowledge, validation checks, failure branches,
transfer notes, and source boundaries without requiring an MCP server or a
runnable paper codebase.

\section{Method}

PaperToSkill represents each converted paper as a skill package. The main file
is \texttt{SKILL.md}; optional references include a source map that links
generated skill bullets to source note sections and line spans. The schema
contains identity, central contribution, use conditions, required artifacts,
workflow steps, validation checks, failure cases, transfer notes, and source
notes that distinguish source-backed material from inferred guidance.

The deterministic extractor reads a source-anchored paper note, normalizes
Markdown sections and list items, collects candidates from abstract, methods,
experiments, limitations, and transfer sections, then emits a compact skill.
Candidate limits keep the artifact within a practical context budget while
preserving method, validation, and failure coverage. The extractor also writes a
\texttt{references/source\_map.json} file so later audits can inspect which
source section supports each generated instruction.

The text-to-note scaffold is an earlier preprocessing step. It operates on
newline-numbered extracted text, detects broad paper regions, selects line
windows for methods, experiments, and limitations, and preserves line anchors.
For two-column extracted text, it keeps raw line spacing and selects the
keyword-bearing column while retaining the original source line range. The
output is explicitly marked as an audit scaffold.

\section{Experimental Setup}

We evaluate four real agent-method papers: AI Scientist-v2, Reflexion, AIDE,
and Toolformer. For each paper, we create a curated source-anchored note from
the extracted paper text, generate a PaperToSkill skill, and compare it against
two context baselines: a generic summary and an abstract-only context. For
transfer ablation, we compare the full skill against the same skill with
\texttt{Transfer Notes} removed and against the generic summary.

The metrics are deterministic: skill rubric score, context coverage,
source-span validation, offline transfer-readiness, word count under a
1200-word compactness budget, character-based token/cost proxy, and a local
tokenizer-aware input proxy using \texttt{o200k\_base}. For saved
model-ablation responses, we additionally compute a local output-token proxy.
These metrics are reproducible gates. They do not replace live agent execution,
real billing records, invoice evidence, success-per-dollar accounting, or
completed human fidelity annotation. The current package uses local token
accounting over input-token and saved-response output-token proxies. We also
prepare model-ablation prompt packets for Claude Opus 4.8, a GPT-family slot
requested as GPT 5.5, and a DeepSeek slot, and include a protocol-aware runner
and response evaluator. The current two-case protocol has 6/6 saved and scored
rows across Claude, GPT-family, and DeepSeek slots. This is saved-response
output-contract evidence only; it is not live downstream task success,
provider billing, or a broad model-quality result. Provider/model availability
failures are logged separately from model-quality results.

\input{papertoskill_tables}

\section{Results}

All four generated skills score 20/20 on deterministic skill rubrics. On context
coverage, the generated skills score 7.867/9 for AI Scientist-v2, 8.267/9 for
Reflexion, 9.1/10 for AIDE, and 8.9/10 for Toolformer. The generic-summary and
abstract-only baselines score substantially lower across all four cases
(Table~\ref{tab:main-results}).

The generated skills remain compact: 782 words for AI Scientist-v2, 479 words
for Reflexion, 927 words for AIDE, and 943 words for Toolformer, all under the
1200-word budget. Source validation finds support rates of 0.938, 1.0, 1.0,
and 1.0 with zero invalid line ranges. The one weak AI Scientist-v2 span is
recorded as a boundary case rather than treated as fully verified.

The tokenizer-aware context proxy shows the same compactness pattern relative
to full paper contexts. Under \texttt{o200k\_base}, generated skills use between
2.39\% and 9.65\% of the input tokens required by full extracted paper text
(Table~\ref{tab:cost-proxy}). A separate character-proxy report preserves the
earlier $\lceil\mathrm{characters}/4\rceil$ estimate for reproducibility. We
interpret both as coverage-preserving compression relative to full-paper
context, not as provider billing evidence or a claim that skills are the
shortest possible context.

Transfer-note ablations show a consistent offline readiness pattern. Full
skills score 10/10 on the readiness metric across all four curated papers.
Removing \texttt{Transfer Notes} lowers readiness to 7.6/10 in all four cases,
while generic summaries score between 1.2/10 and 2.25/10. This supports the
narrower claim that transfer notes encode artifact-readiness signals. It does
not yet prove improved live cross-harness success.

The live-transfer saved-response evaluation covers all four paper packets.
AI Scientist-v2, Reflexion, AIDE, and Toolformer each have six saved responses
across Codex-style and Claude-style harness prompts and three context variants.
The aggregate evaluator reports 24 total rows, 24 scored rows, 0 pending rows,
and average normalized score 1.0. AI Scientist-v2, Reflexion, and AIDE rows
score 11/11 each; Toolformer rows score 9/9 each. The AIDE run includes one
provider fallback row where \texttt{claude-opus-4-8} closed the connection and
\texttt{claude-opus-4-7} succeeded. These are deterministic output-contract
scores over saved response files, not human-rated semantic fidelity or
downstream execution-outcome rates.

The automatic note scaffold is evaluated separately on Toolformer and AIDE. The
Toolformer auto-note-derived skill scores 20/20 on the deterministic rubric,
9.3/10 on context coverage, 10/10 on offline transfer readiness, and 1.0
source-span support with zero invalid ranges. The AIDE auto-note-derived skill
scores 20/20, 8.467/10 on context coverage, 9.5/10 on offline transfer
readiness, and 1.0 source-span support with zero invalid ranges
(Table~\ref{tab:auto-note}). These results support the narrower claim that
extracted text can seed auditable note scaffolds. The local pipeline also
includes a direct-PDF smoke path through \texttt{pdftotext -layout} that records
the generated text source in the manifest, but this does not establish robust
arbitrary-PDF automation.

We additionally compare PaperToSkill against Paper2Agent at the artifact and
workflow level using the Paper2Agent paper text and current PaperToSkill
artifacts. The bounded comparison covers required inputs, generated artifact
type, setup burden, validation checks, failure handling, source traceability,
and runtime dependency. All seven criteria are source-backed and ready in the
local comparison table. The result clarifies positioning: Paper2Agent produces
MCP servers for papers with associated codebases, while PaperToSkill produces
portable natural-language skills with explicit source and failure boundaries.
This comparison does not run Paper2Agent, deploy an MCP server, or establish an
end-to-end baseline performance result.

\section{Usage Examples and Model Ablations}

The repository includes usage examples for three practical workflows:
loading a generated skill in a Codex-style harness, generating an automatic
note scaffold from extracted text or a smoke-tested PDF source, and building
model-ablation prompts. The
model-ablation protocol produces a prompt grid for Claude Opus 4.8, a
GPT-family slot requested as GPT 5.5, and a DeepSeek slot. In the current
protocol, the saved-response evaluator reports 6 total rows, 6 scored rows, 0
pending rows, and average normalized score 1.0. The GPT-family protocol refresh
completed both rows with \texttt{gpt-5.5}. The DeepSeek run completed both rows
with \texttt{deepseek-v4-flash}. The latest Claude protocol refresh used the
correct Anthropic Messages path but was blocked by provider HTTP 502; the
scored Claude rows come from previously saved response files. A local
output-token proxy over all six saved model-ablation responses counts 9,594
\texttt{o200k\_base} output tokens. This 6/6 saved-response score is not live
downstream task success, provider billing, live-invoice, success-per-dollar
evidence, or a broad model-quality result.

The bounded AI-Scientist-v2 integration path is complete for the local marker
smoke and one full live run. The smoke report records a complete marker-contract
response, and the full-run handoff records one completion directory. The
AI-Scientist-v2-generated synthetic benchmark ties skill and full-excerpt
task-success rates at 0.80 while the skill uses fewer tokens (86.2 versus
113.2); abstract, generic-summary, and no-context baselines score 0.20, 0.00,
and 0.00. A retrieval-depth sensitivity branch reaches 1.00 only when K=all.
This is bounded integration and synthetic sensitivity evidence, not human
fidelity, real-data validation, or broad live research-task success.

\section{Discussion}

The main result is not that PaperToSkill ``understands'' papers in the broadest
sense. The result is that a structured paper-note-to-skill pipeline can preserve
operational knowledge that summaries discard. The skills include workflow
steps, validation checks, failure cases, and source anchors, which are exactly
the parts needed when a user wants an agent to apply a paper rather than merely
explain it.

The benchmark also reveals useful failure modes. During the AIDE case, the
initial extractor capped workflow, validation, and failure candidates too
aggressively, which dropped data-preview and LLM-cost content. Earlier phases
also exposed title parsing, source-audit mapping, and source-span line-counting
bugs. These records are the type of failure branches PaperToSkill should
preserve: not just final successes, but the ways extraction, evaluation, or
external dependencies can lose important operational content.

\section{Limitations}

The current work has several important limits. First, the main benchmark
converts curated notes rather than arbitrary PDFs; automatic note scaffolds have
only been retained on Toolformer and AIDE extracted text. Second, the metrics
are deterministic and partly lexical, so they can over-credit exact matches or
under-credit valid paraphrases. Third, live-transfer saved-response coverage is
complete for the current prompt-packet protocol, but the scorer is
deterministic and contract-based rather than a human semantic-fidelity or live
execution-outcome evaluation. Fourth, the bounded AI-Scientist-v2 smoke and
full live run is complete, but its positive result is synthetic and should not
be read as broad live task success. Fifth,
human-fidelity packets, a reviewer handoff guide, a stricter blank annotation
template, and a summary script are prepared, but no independent annotations have
been completed. Sixth,
compactness is measured by word count, deterministic character proxy, local
tokenizer-aware input-token proxy, and local output-token proxy for saved model
responses, not by provider-specific prices, live invoices, or live success per
dollar.

\section{Conclusion}

PaperToSkill turns paper-derived procedural knowledge into portable,
human-editable skills. In a four-paper benchmark, generated skills are compact,
source-grounded, structurally valid, and more operationally complete than short
summary baselines under deterministic evaluation. The next stage is to add
human fidelity review, evaluate downstream execution outcomes separately from
saved-response scoring,
extend the bounded Paper2Agent artifact comparison into a real executable MCP
baseline if resources permit, and test papers whose contributions are less
naturally procedural.

\bibliography{papertoskill_refs}

\end{document}
